### 2012-06-13 trunk/ijcis-pos-model/sources/preliminaries.tex
version sent to Olga and Nicolas

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} &\\
\nonumber
Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} & \\
\nonumber
Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}

Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} :  \mu_M(m)=1 
\end{eqnarray}
\end{definition}

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Trapezoidal membership function.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Trapezoidal membership function.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

The most convenient form of the membership function of a fuzzy number is a triangular form. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}

The membership function of such a fuzzy number is then given by:

@@

\begin{align}
\mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
\begin{cases}
\nonumber
0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
\frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
-\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
\end{cases}
\end{align}

\subsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.

The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
\end{align}

So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint. However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

\begin{example}
Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:

@@

There is a special boolean combination of constraints that is of particular interest:
\begin{definition}
\emph{Convex combination of ill-known constraints}. Consider the $C_1$ and $C_2$ two ill-known constraints. A convex combination of the constraints is given by:
\begin{align}
\label{eq:convex-combination}
CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
\end{align}
In a more general way, it is possible to define the convex combination of an n-ary vector  $\mathbf{C}$ of constraints:
\begin{align}
\label{eq:nary-convex-combination}
CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
\end{align}
\end{definition}

\begin{theorem}
\label{th:convex-combination-ill-known-constraints}
Consider the convex combination $CC$ of any n-ary vector $\mathbf{C} = \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$. Then, if $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex, then  $\pi_{CC}$ is also convex.
\end{theorem}

\begin{proof}
Let $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$. Then:
\begin{align}
\label{eq:proof-convex1}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) = \\
\min  \big( \pi_{C_1(Z_1)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right), \ldots,\\
   \pi_{C_n(Z_n)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right)  \big)
\end{align}
Since $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex:
\begin{align}
\label{eq:proof-convex2}
\pi_{C_1(Z_1)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
\nonumber
\min \big(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right)  \big)\\
\nonumber
\vdots \\
\nonumber
\pi_{C_n(Z_n)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
\nonumber
\min \big(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \big) 
\end{align}
Then, by using equation \eqref{eq:proof-convex1}:
\begin{align}
\label{eq:proof-convex3}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right) \right),\\
\nonumber
\ldots, \min \left(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
\end{align}
Which is equivalent to the following:
\begin{align}
\label{eq:proof-convex4}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right), \ldots, \pi_{C_n(Z_n)} \left( x_1 \right) \right),\\
\nonumber
\ldots, \min \left(\pi_{C_1(Z_1)} \left( x_2 \right), \ldots, \pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
\end{align}
Finally we obtain:
\begin{align}
\label{eq:proof-convex5}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big(\pi_{CC} \left( x_1 \right), \pi_{CC} \left( x_2 \right) \big)
\end{align}
\end{proof}

@@

\begin{definition}
Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$. The value $X$ is defined by means of the convex combination:
\begin{align}
\label{eq:ill-known-value-by-convex-constraints}
X \triangleq CC 
\end{align}

The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align} 
Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demostrated in theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

>>>>

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$; the proposition can be seen as a possibilistic variable. The possibility and necessity that $p$ = $T$ (the proposition holds) demand more attention. This possibility and necessity is noted here as:
\begin{align}
\label{propholdsposs}
\text{Possibility that $p$ = $T$ (p holds):} &\\
\nonumber
Pos(p) = \pi_p(T)  \\
\label{propholdsnecc}
\text{Necessity that $p$ = $T$ (p holds):} & \\
\nonumber
Nec(p) = 1-\pi_p(F) 
\end{align}

Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, while equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work deals with ill-known intervals. These are ill-known sets, defined and represented via a starting and ending points, which are ill-known values. The elements of the set are the values between the starting and ending points. A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right]  \\ 
\nonumber
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
\nonumber
\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\nonumber
\exists m \in \mathbb{R} :  \mu_M(m)=1 
\end{eqnarray}
\end{definition}

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}). It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. The values $\left[\alpha, \beta, \gamma, \delta\right]$ represent a trapezoidal possibility distribution defined by $\mu_T$:

% Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as . The corresponding membership function definition for this  is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
\nonumber
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

The representation for a triangular possibility distribution is given by three values $\left[D,a,b \right]$. The distribution is obtained from a trapezoidal possibility distribution by replacing the values $\left[D-a, D, D, D+b \right]$.

% The most convenient form of the membership function of a fuzzy number is a triangular form. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
% \begin{itemize}
% \item
% $D$ denotes the single value in the core of $M$
% \item
% $D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
% \item
% $D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
% \end{itemize}
% 
% The membership function of such a fuzzy number is then given by:
% 
% \vspace{-10pt}
% 
% \begin{align}
% \mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
% \begin{cases}
% \nonumber
% 0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
% \frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
% -\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
% \end{cases}
% \end{align}

\subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
In our case, we need to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. To compute that, we introduce the concept of \emph{ill-known constraint}\cite{Pons2011}.

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
\label{eq:ill-known-constraint}
C \triangleq (R,V)
\end{align}
Set $A$ satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

An example of an ill-known constraint is $C_{ex} = (<, X)$. Some set $A$ satisfies $C_{ex}$ if $\forall a \in A : a < X$, given the possibilistic variable $X$.

The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be computed. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

@@

\begin{align}
\label{ill-known-pos}\
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \\
\label{ill-known-nec}
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) 
\end{align}

So far, we have shown how it can be verified whether a set satisfies or not an ill-known constraint. The interval evaluation problem is explained in a more general context in \cite{Pons2011}.

\begin{example}
As example, consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint. Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively. (See Figure \ref{fig:example-ill-known}).

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey. }
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
\end{align}

\end{example}

It is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation (explained earlier) requires that all elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:\lambda(A)\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only those connectives will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1 \ldots n}\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

\begin{example}
Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.

@@

There is a special boolean combination of constraints that is of particular interest; let us see it.
\begin{definition}
\emph{Conjunctive combination of ill-known constraints}. Consider a universe $U$. Let $R_1, R_2 \subseteq U$ be two binary relations in U. Let $X_1, X_2$ be two fixed ill-known values in $U$. Let  $C_1 = (R_1, X_1)$ and $C_2 = (R_2, X_2)$ be two ill-known constraints. A conjunctive combination of both constraints is given by:
\begin{align}
\label{eq:convex-combination}
CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
\end{align}
In a more general way, it is possible to define the conjunctive combination of an n-ary vector of constraints:
\begin{align}
\label{eq:nary-convex-combination}
CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
\end{align}
\end{definition}

\begin{theorem}
\label{th:convex-combination-ill-known-constraints}
Consider the conjunctive combination $CC$ of any n-ary vector $\left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$. Then, if $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex, then  $\pi_{CC}$ is also convex.
\end{theorem}

\begin{proof}
Let $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$. Then:
\begin{align}
\label{eq:proof-convex1}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) = \\
\nonumber
\min  \big( \pi_{C_1(Z_1)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right), \ldots,\\
\nonumber
   \pi_{C_n(Z_n)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right)  \big)
\end{align}
Since $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex:
\begin{align}
\label{eq:proof-convex2}
\pi_{C_1(Z_1)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
\nonumber
\min \big(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right)  \big)\\
\nonumber
\vdots \\
\nonumber
\pi_{C_n(Z_n)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
\nonumber
\min \big(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \big) 
\end{align}
Then, by using equation \eqref{eq:proof-convex1}:
\begin{align}
\label{eq:proof-convex3}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right) \right),\\
\nonumber
\ldots, \min \left(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
\end{align}
Which is equivalent to the following:
\begin{align}
\label{eq:proof-convex4}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right), \ldots, \pi_{C_n(Z_n)} \left( x_1 \right) \right),\\
\nonumber
\ldots, \min \left(\pi_{C_1(Z_1)} \left( x_2 \right), \ldots, \pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
\end{align}
Finally we obtain:
\begin{align}
\label{eq:proof-convex5}
\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
\nonumber
\min \big(\pi_{CC} \left( x_1 \right), \pi_{CC} \left( x_2 \right) \big)
\end{align}
\end{proof}

@@

\begin{definition}
\label{def:convex-combination-ill-known-constraints}
\emph{Ill-known value defined by conjunctive combination of constraints.}
Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$. Then, the value $X$ is defined by means of the convex combination:
\begin{align}
\label{eq:ill-known-value-by-convex-constraints}
X \triangleq CC 
\end{align}

The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align} 
Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

\begin{example}
As an example, consider the hospital database. When a patient enters o leaves a medical service in hospital, a record tracking the patient is stored. The exact time for this is not precisely known. Consider now that $X = \left[12, 2, 2\right]$ is the time in hours when a patient left the emergency service and $Y = \left[18, 2, 1 \right]$ is the time in hours when the patient left the hospital. It is possible to obtain $Z$ (the period of time between the emergency service and the left of the patient) by using a conjunctive combination of constraints.
